---
title: "Mini Project 2"
author: "Sam English"
date: 10/02/2024
---

```{r, message = FALSE, warning = FALSE}

tuesdata <- tidytuesdayR::tt_load('2021-04-20')
tuesdata <- tidytuesdayR::tt_load(2021, week = 17)

netflix <- tuesdata$netflix

netflix

```

```{r, message = FALSE, warning = FALSE}

library(dplyr)
library(tidytext)
library(tidyr)
library(stringr)
library(ggplot2)

```

```{r, message = FALSE, warning = FALSE}
stop_words <- c("the", "a", "of", "and", "in", "on", "with", "is", "to", "for", "at", "s", "2", "i")

common_words <- netflix |>
  mutate(title = str_to_lower(title),                               
         title = str_replace_all(title, "[[:punct:]]", " ")) |>     
  rowwise() |>                                                      
  mutate(words = str_split(title, "\\s+")) |>                       
  unnest(words) |>                                                  
  filter(!words %in% stop_words,                                     
         words != "") |>                                            
  group_by(words) |>                                      
  mutate(n = n()) |>                                            
  arrange(desc(n)) |>                                          
  distinct(words, n, .keep_all = TRUE) |>
  select(words, n) |>
  head(10)

common_words

```

```{r, message = FALSE, warning = FALSE}

ggplot(common_words, aes(x = reorder(words, -n),  y = n)) + 
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Most Common Words in Netflix Titles", 
       x = "Words", 
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```
This plot shows the most common occurring words in Netflix titles, if you remove stock words such as "the", "and" and "of". Some obvious 'buzzwords' include "love, "story" and "movie". Some key words that surprised me include "christmas", but it also makes sense as Christmas movies, much like songs, are very popular at that specific time of year. "Man" is also an interesting observation considering there is no appearance of "woman" in the top 10 words. 

```{r, message = FALSE, warning = FALSE}

netflix_by_year <- netflix |>
  group_by(release_year, type) |>
  summarise(count = n())  

netflix_by_year

ggplot(netflix_by_year, aes(x = release_year, y = count, color = type)) +
  geom_line() +
  labs(title = "Number of Movies and TV Shows Released by Year on Netflix", 
       x = "Release Year", 
       y = "Count") +
  theme_minimal()

```
This graph shows the number of movies and TV shows and their corresponding release years. It shows the increase in number of movies as the year heads towards present day. It also demonstrates the disparity of movies vs. TV shows. And that there has always been more movies than TT shows on Netflix. An interesting observation is in 2021, where there is growth in the number of TV shows released in this year, and a decline in movies. It is also cool to note that the earliest title on Netflix was released in 1925, and it is a TV show!! 

```{r, message = FALSE, warning = FALSE}

total_titles <- nrow(netflix)

prop_titles <- netflix |>
  mutate(title = str_to_lower(title)) |> 
  summarise(
    count_numbers = sum(str_detect(title, "\\d")),  
    count_the_anywhere = sum(str_detect(title, "the")),  
    count_the_start = sum(str_detect(title, "^the"))  
  ) |>
  mutate(
    The_Anywhere = count_the_anywhere / total_titles, 
    Start_With_The = count_the_start / total_titles, 
    Digit_Anywhere = count_numbers / total_titles  
  )

prop_titles

prop_titles_long <- prop_titles |>
  pivot_longer(
    cols = c(The_Anywhere, Start_With_The, Digit_Anywhere),  
    names_to = "condition",              
    values_to = "proportion"             
  ) |>
  select(condition, proportion)

prop_titles_long 

ggplot(prop_titles_long, aes(x = condition, y = proportion, fill = condition)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Proportions of Titles in Netflix",
    x = "Title Condition",
    y = "Proportion",
    fill = "Condition"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
This plot shows the proportions of titles that start with "the", contain "the" or contain any digit. Almost exactly twice as many titles contain "the" anywhere in the title than titles that start with "the". Almost a quarter of Netflix titles contain the word "the". Far fewer titles contain digits, at only 5.5%.   


