---
title: "Netflix Title Analysis"
author: "Sam English"
date: 10/02/2024
---

This is my second project using R to visualize data. The focus in this project is to use piping to organize data so it is easier to understand and present. I focused on data involving Netflix titles and made three visualizations. Firstly, I found the most common words in titles of Netflix Movies and TV Shows, removing filler words like 'the' and 'and'. Next, I compared the number of titles on Netflix that were movies or TV Shows, and their release dates. Finally, I displayed the percent of titles that contain a digit anywhere in the title, start with "the" or have the word "the" anywhere and showed this information on a bar graph.   


This analysis uses data from the tidy tuesday data source on titles of Netflix movies and TV show. [Netflix Titles]
<https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-04-20/readme.md>

the original data can be found on Kaggle at <https://www.kaggle.com/datasets/shivamb/netflix-shows>. It was published by Bansal. The data is titled "Netflix Movies and TV Shows", and consists of listings of all the movies and tv shows available on Netflix, along with details such as - cast, directors, ratings, release year, duration, etc.



Here is the entire dataset that I used. 
```{r, message = FALSE, warning = FALSE}

tuesdata <- tidytuesdayR::tt_load('2021-04-20')
tuesdata <- tidytuesdayR::tt_load(2021, week = 17)
netflix <- tuesdata$netflix
netflix
```

```{r, message = FALSE, warning = FALSE}

library(dplyr)
library(tidytext)
library(tidyr)
library(stringr)
library(ggplot2)

```

```{r, message = FALSE, warning = FALSE}
stop_words <- c("the", "a", "of", "and", "in", "on", "with", "is", "to", "for", "at", "s", "2", "i")

common_words <- netflix |>
  mutate(title = str_to_lower(title),                               
         title = str_replace_all(title, "\\p{P}", " ")) |>     
  rowwise() |>     #Process each row of the dataset separately, allowing for operations like splitting words in each title.                                                  
  mutate(words = str_split(title, "\\s+")) |> # split words separated by spaces             
  unnest(words) |>      # Each word from a title becomes its own row.             
  filter(!words %in% stop_words, words != "") |>                                            
  group_by(words) |>                                      
  mutate(n = n()) |>                                            
  arrange(desc(n)) |>                                          
  distinct(words, n, .keep_all = TRUE) |>
  select(words, n) |>
  head(10)

common_words

```

```{r, fig.alt= 'Bar chart showing the most common words in Netflix titles, with ‘love’ being the most frequent, followed by ‘my’ and ‘you.’ Frequencies decrease for words like ‘man,’ ‘Christmas,’ ‘world,’ and others.', message = FALSE, warning = FALSE}

ggplot(common_words, aes(x = reorder(words, -n),  y = n)) + 
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Most Common Words in Netflix Titles", 
       x = "Words", 
       y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```
This plot shows the most common occurring words in Netflix titles if you remove stock words such as "the", "and" and "of". Some obvious 'buzzwords' include "love, "story" and "movie". Some key words that surprised me include "Christmas", but it also makes sense as Christmas movies, much like songs, are very popular at that specific time of year. "Man" is also an interesting observation considering there is no appearance of "woman" in the top 10 words. 

```{r, fig.alt= "Line chart showing the number of movies and TV shows released by year on Netflix. The red line represents movies, peaking around 2019, and the blue line represents TV shows, with a similar peak slightly after movies. Both lines show minimal activity before 2000, followed by a sharp increase post-2010.", message = FALSE, warning = FALSE}

netflix_by_year <- netflix |>
  group_by(release_year, type) |>
  summarise(count = n())  

netflix_by_year

ggplot(netflix_by_year, aes(x = release_year, y = count, color = type)) +
  geom_line() +
  labs(title = "Number of Movies and TV Shows Released by Year on Netflix", 
       x = "Release Year", 
       y = "Count") +
  theme_minimal()

```
This graph shows the number of movies and TV shows and their corresponding release years. It shows the increase in number of movies as the year heads towards present day. It also demonstrates the disparity of movies vs. TV shows. And that there has always been more movies than TT shows on Netflix. An interesting observation is in 2021, where there is growth in the number of TV shows released in this year, and a decline in movies. It is also cool to note that the earliest title on Netflix was released in 1925, and it is a TV show!! 

```{r, fig.alt= "Bar chart titled “Proportions of Words in Netflix Titles” showing three conditions: “Digit” anywhere (red, around 5%), titles starting with “The” (green, around 15%), and “The” appearing anywhere (blue, around 25%). The proportions increase from left to right.", message = FALSE, warning = FALSE}

total_titles <- nrow(netflix)

prop_titles <- netflix |>
  mutate(title = str_to_lower(title)) |> 
  summarise(
    count_numbers = sum(str_detect(title, "\\d")),  
    count_the_anywhere = sum(str_detect(title, "the")),  
    count_the_start = sum(str_detect(title, "^the"))) |>
  mutate(
    The_Anywhere = count_the_anywhere / total_titles, 
    Start_With_The = count_the_start / total_titles, 
    Digit_Anywhere = count_numbers / total_titles  
  )

prop_titles

prop_titles_long <- prop_titles |>
  pivot_longer(
    cols = c(The_Anywhere, Start_With_The, Digit_Anywhere),  
    names_to = "condition",              
    values_to = "proportion"             
  ) |>
  select(condition, proportion)

prop_titles_long 

ggplot(prop_titles_long, aes(x = condition, y = proportion, fill = condition)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Proportions of Words in Netflix Titles",
    x = "Title Condition",
    y = "Proportion",
    fill = "Condition"
  ) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(labels = c(
    The_Anywhere = '"The" anywhere',
    Start_With_The = 'Start with "The"',
    Digit_Anywhere = '"Digit" anywhere')) +
  scale_fill_discrete(labels = c(
    The_Anywhere = '"The" anywhere',
    Start_With_The = 'Start with "The"',
    Digit_Anywhere = '"Digit" anywhere')) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
This plot shows the proportions of titles that start with "the", contain "the" or contain any digit. Almost exactly twice as many titles contain "the" anywhere in the title than titles that start with "the". Almost a quarter of Netflix titles contain the word "the". Far fewer titles contain digits, at only 5.5%. I expected this proportion to be higher.    


